Extracting Angular Observables without a Likelihood\\and Applications to Rare Decays

### 2015-03-12 draft.tex
Update with Fred

Our initial motivation for studying what we wish to call the
\emph{method of moments} is the determination of angular observables
in the rare FCNC-mediated decay $\bar{B}\to \bar{K}^*(\to
\bar{K}\pi)\ell^+\ell^-$. However, the method we describe in the
following is general and applies to all decay or scattering
processes that can be formulated in terms of an orthogonal basis of angular functions.
We find a previous work \cite{Dighe:1998vk} that
advocates this method chiefly for the determination of angular
observables in non-leptonic $B$ decays but also mentions the
applicability to semileptonic decays. Our aim is to improve upon this
previous work by studying the uncertainties that are introduced by
mismodeling of the angular distribution, and by working out a recipe
to determine and unfold detector effects. The latter is crucial for
the application of the method to real data. We show that
the method of moments has several major advantages over the usual approach
based on likelihood fits:
\begin{enumerate}
    \item Likelihood fits have convergence problems for a small number of
        events, and can require reparametrizations and/or approximations
        for a successful fit to the signal PDF. As an example see LHCb measurement of $\bar{B}\to \bar{K}^*(\to \bar{K}\pi)\mu^+\mu^-$\cite{Aaij:2013qta}. The method of moments does not require reparametrizations or approximations.

    \item Likelihood fits  can be unstable in case the underlying
        physical model is only partially known. This can lead to
        overestimating the number of physical parameters, and consequently
        inhibits the convergence of the fits. As an example, this type
        of problem occured in toy studies of the decay $B\to K^*(\to K\pi)\ell^+\ell^-$
        as reported in \cite{Egede:1142152}. It was subsequently solved
        when a missing symmetry relation between the angular observables
        was found and applied, thereby reducing the number of fit parameters.
    \item Mismodeling the underlying physics model can result in systematic bias
        in likelihood fits. We will
        show that the method of moments is insensitive to a certain type of mismodeling;
        i.e. introducing a cutoff in a partial-wave expansion of the signal PDF.

@@

\fred{For us, the major advantage is that the experiment can obtain all angular observables at once, and if there are more, one doesn't need more data. I'm not sure this becomes clear from the text, point 1.}\marcin{I think it is ok}

We continue with basic definitions that pertain to angular observables, and our
results in the subsequent sections. Let $\thvec$ denote the set of all
angles, and let $\nuvec$ denote the set of all
other non-angular kinematic variables needed to fully specify the
final state of the process under study. For example, $\nuvec$ may
include invariant masses or center-of-mass energies. We define an
angular observable $S_i$ as a coefficient in the probability density
function (PDF), $P(\nuvec, \thvec)$, of the process by
\begin{align}
    \label{eq:def-P}
    P(\nuvec, \thvec) \equiv \sum_i S_i(\nuvec) \times f_i(\thvec)\,.
\end{align}
Here, the dependence on the decay angles $\thvec$ has been
explicitly factored out in terms of the angular functions
$\{f_i(\thvec)\}$. We assume there exists a dual basis of functions
$\{\dual{f}_i(\thvec)\}$ such that the orthonormality relations
\begin{equation}
    \label{eq:def-ortho-rel}
    \int_\Omega \rmdx{\vec{\theta}} \dual{f}_i(\thvec) f_j(\thvec)  = \delta_{ij}
\end{equation}
hold with $\Omega$ representing the full angular phase space relevant
to the process.  For particle decays, $P$ is generally expressed in
terms of the fully differential decay width,
\begin{align}
    \label{eq:def-P-decay}
    P(\nuvec, \thvec) \equiv \frac{1}{\Gamma}\frac{\rmdx{^{2}\Gamma}}{\rmdx{\vec\nu} \rmdx{\vec\theta}}\,,
\end{align}
where $\Gamma$ is the total decay width. For a scattering process, one can similarly use
\begin{align}
    \label{eq:def-P-scattering}
    P(\nuvec, \thvec) \equiv \frac{1}{\sigma}\frac{\rmdx{^{2}\sigma}}{\rmdx{\vec\nu} \rmdx{\vec\theta}}\,,
\end{align}
where the total cross section $\sigma$ is used for the
normalization. Since the determination of the total decay width or
total cross section can be quite difficult, we emphasize that
different normalizations for $P$ can be used.  For instance, the total
decay width (or cross section) of the process of interest can be
replaced by the corresponding quantity of a control-channel
process. This change of normalization is equivalent to a linear rescaling of
the angular observables $\lbrace S_i\rbrace$; thus ratios or similar suitable combinations
of the angular observables are not affected by \add{a} change of normalization.\\

Our method is an extension of the classical method of moments with
orthogonal functions~\cite[sec. 8.2]{James:2006zz}. The only difference
is that conventionally the angular functions are assumed
\emph{self-dual}, $\dual{f}_i = f_i$. However, it suffices that the
system of angular functions $\{f_i(\thvec)\}$ \emph{can} be transformed into
an orthonormal basis. \add{We find it convenient to work
in the basis of Legendre polynomials, which is not self-dual.
However, the statistical properties of the method of moments are equivalent for both self-dual and non-self-dual
bases, provided that one replace $f_i \to \dual{f}_i$ appropriately.}
\fred{I'm confused. What are the two approaches? Do we consider Legendre polynomials as an orthonormal basis or not?}
Using the ansatz
\begin{equation}
  \label{eq:dual-ansatz}
  \dual{f}_i = \sum_{j} a_{ij} f_j \,,
\end{equation}
the dual basis needs to be worked out case by case through solving the
linear system of equations (\ref{eq:def-ortho-rel}). For a selection
of hadron decays with a $b$ quark in the initial state and two leptons
in the final state, we list the dual bases in a series of appendices
\ref{app:btokll} through \ref{app:btokstarll}. Note that a
similar analysis was done in~\cite{Dighe:1998vk} for the decays $B \to
J/\psi \phi$ and $B \to J/\psi K^{*}$.

@@

Integration over the non-angular variables yields
\begin{equation}
    \langle S_i\rangle
    \equiv \int \rmdx{\vec\nu} S_i(\nuvec)
    = \int \rmdx{\vec \nu} \left[\int_{\Omega} \rmdx{\vec \theta} P(\nuvec,\thvec) \dual{f}_i(\thvec) \right].
\end{equation}
\fred{Why is there $\Omega$ but nothing for $\nuvec$?}
\danny{a) intoduced $\Omega$ one equation earlier, b) we always need
and integration of the \emph{entire} angular phase space; the integration
over the non-angular phase space is irrelevant to the discussion.}
The remainder of this section describes the method of moments, in
which we replace the analytical integration by Monte Carlo (MC)
estimates.  The central tenet of MC integration is the fact that the
expectation value $E_P[g]$ of some function $g(x)$ under the
probability density $P(x)$,
\begin{equation}
    E_P[g] \equiv \int \rmdx{x} P(x) g(x),
\end{equation}
can be approximated by the consistent and unbiased
estimator $\est{E_P[g]}$~\cite[sec. 8.2]{James:2006zz}
\begin{equation}
    \label{eq:mc-id}
    E_P[g] \to \widehat{E_P[g]} \equiv \frac{1}{N} \sum_{n=1}^N g(x^{(n)}) \,,\,    x^{(n)} \sim P
\end{equation}
due to the strong law of large numbers for $N \to \infty$, assuming
that the variates $x^{(n)}$, $n = 1, \dots, N$, are distributed as
$P$.
Throughout this letter we denote all MC estimators with a wide hat.\\

@@

\item The estimate is unbiased for any $N$. In contrast, the
  maximum-likelihood estimate has a bias of order
  $1/N$~\cite{Cox:1968}. In practice, one should keep in mind the
  bias-variance trade-off: it is a well known phenomenon that removing
  the bias usually leads to an increase in variance of the sampling
  distribution of the estimator and vice versa~\cite[sec. 7.3]{James:2006zz}. From a
  Bayesian decision-theory point of view, both contribute similarly to
  the expected loss associated with deciding on just one value of the
  unknown parameter. One should therefore not prefer the method of moments
  over likelihood fits just because the former reduces
  the bias~\cite[sections 13.8,17.2]{jaynes:2003} \marcin{we said we cahnge this but after reading it again I like it and would not change}. In fact, for
  the results discussed below in \refsec{numerics}, the likelihood
  fits --- if they converge --- produce uncertainties (10--15)\%
      smaller than those from the method of moments.
\item The approximate multivariate Gaussian distribution of $\vecest{S}$
    allows easier and more \add{precise} transfer of the information in the
  data to interested theorists for more accurate fits of
  standard-model and new-physics parameters \cite{Altmannshofer:2013foa,
Descotes-Genon:2013wba,Beaujean:2013soa},
  or for more precise predictions of optimized observables; see e.g.,
  \cite{Egede:2008uy,Egede:2010zc,Bobeth:2010wg,Becirevic:2011bp,
    Bobeth:2012vn,Matias:2012xw,DescotesGenon:2012zf} for definitions
  of such optimized observables in $B\to K^*\ell^+\ell^-$ decays,
  \cite{Faller:2013dwa} for application to the decay $B\to
  \pi\pi\ell^-\bar\nu_\ell$, and \cite{Boeer:2014xx} for observables
  in $\Lambda_b\to\Lambda(\to N\pi)\ell^+\ell^-$). While the
  likelihood also approaches a multivariate Gaussian as $N \to
  \infty$, the two methods differ in their utility as input for
  theorists if $\widehat{\vec{S}}$ is not well
  inside the physical region. For example, suppose there are two
  angular observables that are constrained to a triangular region by
  phase-space or unitarity arguments as
  \begin{equation}
    \label{eq:constr-ex}
    |S_1| + S_2 \le 1, \, S_1 \in [-1,1],\, S_2 \in [0,1] \,.
  \end{equation}
  It may (and often does) happen in practice that $\widehat{\vec{S}}$
  is close or even outside the allowed region such that a significant
  part of the probability mass covers unphysical values. In a Bayesian
  fit, one would take $\mathcal{N}(\vecest{S} | \vec{S},
  \widehat{\Sigma})$ as the sampling distribution of the ``data'', and
  simply set a uniform prior on the triangle in the $\vec{S}$ plane
  defined by \refeq{constr-ex} to have a well defined problem. This
  could be trivially combined with other independent information in a
  global fit. Someone with a different physics model might have to
  consider a
  different physical region, and could incorporate it just as easily.\\

  For a likelihood fit, the constraint needs to be part of the
  analysis performed by the experimental collaboration, and the
  resulting likelihood as a function of $\vec{S}$ is distinctly not
  Gaussian.  Communicating such a result has
  proved to be challenging due to technical reasons (such as data formats,
  size etc.) %\add{as well as political reasons; i.e., the publication of
 % the likelihood also publishes collaboration-internal data.}
  This leads to the
  undesirable situation that only the mode and standard errors
  are reported, and theorists often include the results as independent
  measurements with a Gaussian distribution and disregard the boundary
  problem as well as correlations altogether.

\end{enumerate}

@@

For the selected class of decays, we can describe that problem as follows: The PDF $P$ has a fixed dependence on the dilepton
helicity angle $\theta_1$ and the azimuthal angle $\theta_3$. (See also \refapp{btokstarll} for details
on the angular distribution.) However, at the level of decay amplitudes the dimeson system can have an arbitrarily
large total angular momentum $j$; only its third component is restricted to $j_z = -1,0,+1$.
It is then \add{convenient}\del{usual} to compute the angular distribution explicitly in terms of $\theta_1$ and $\theta_3$, but leave
the angular observables dependent on the remaining helicity angle $\theta_2$.
We advocate here that is a sensible procedure to perform an expansion in terms of Legendre polynomials $p_{l}^{(|m|)}$
with respect to the remaining angle $\theta_2$. \add{Here, the angular-momentum indices $l$ and $m$ follows from the usual
rules for addition of the angular momenta $j$ and $\tilde{j}$ of the partial wave expansion of the underlying
amplitude and its complex conjugate: $j - \tilde{j} \leq l \leq j + \tilde{j}$, and $m = j_z + \tilde{j}_z$.}

For the decay at hand, we consider the partial-wave expansion for the angular observables (see \cite{Lee:1992ih} and \refapp{partial-waves})
\begin{equation}
    S_k(\vec{\nu}) \equiv \frac{I_k(\nuvec)}{4\pi \Gamma)}\,,\quad\text{with}\quad \Gamma = I_1(\nuvec) - \frac{I_2(\nuvec)}{3}\,,
\end{equation}
\fred{It not clever to use $k$ as an index and momentum variable}
\danny{I disagree. We use the same notation as in the literature, and people
should be able to distinguish these based on context.} \marcin{I am fine with k}
which reads
\begin{multline}
    \label{eq:expansion}
    S_{k}(\vec{\nu},\cos\theta_2) \\
        \equiv \sum_{j} \frac{1}{n_{l,|m|}} S_{k,l}(\vec{\nu}) p_{l}^{(|m|)}(\cos\theta_2)\sin\theta_2\,.
\end{multline}
The normalization factor $n_{l,|m|}$ is defined in \refeq{legendre-scalar-product}. Within our
example we have (cf. also \refapp{btokstarll} and \cite{Lee:1992ih})
\begin{equation}
    |m| = \begin{cases}
        0\,, & k = 1,2,6\,,\\
        1\,, & k = 4,5,7,8\,,\\
        2\,, & k = 3,9
    \end{cases}\,.
\end{equation}

The angular observables $S_{k,l}$ -- as defined in \refeq{expansion} --
have the merit of a well defined total angular momentum, and thus are physically discriminable.
As a consequence of the orthogonality of the Legendre polynomials, any mismodelling (or rather, lack of modelling of higher
partial-wave observables) \emph{does not affect} the method of moments as discussed in the previous section. That is to say,
ignoring further (orthogonal) terms in the PDF cannot affect $\vecest{S}$ nor its variance.\\
\fred{That's not true. $\vecest{S}$ is enlarged with more angular observables. But the first components are unaffected.}\\

@@

Ascertaining a detector's performance to detect signal events
with accurate determination of the event's angles is generally a
difficult task. In the ideal case, one would have an explicit
probabilistic model of the detector acceptance and could thus
write down the full forward model from which the measured events
arise. In practice, that is not feasible, and one is
forced to simplify the model.
The standard approach is to generate the \emph{true} particle events
$\lbrace x_\text{t}^{(n)}\rbrace =
\lbrace(\vec{\nu}^{(n)}_\text{t},\vec\theta^{(n)}_\text{t})\rbrace$,
$n=1,\dots,N_\text{t}$, from a PDF assumed to describe the bare
physical process, and to propagate those particles through a detailed
simulation of the detector.  The observable traces that the particles
leave in the detector are fed into reconstruction algorithms resulting
in the \emph{detector} events $\lbrace x^{(n)}_\text{d}\rbrace =
\lbrace(\vec{\nu}^{(n)}_\text{d}, \vec\theta^{(n)}_\text{d})\rbrace$,
$n=1,\dots,N_\text{d}$ with $N_\text{d} \le N_\text{t}$.  In general,
the distribution of the detected events is
\begin{equation}
    P_\text{d}(x_\text{d}) = \frac{1}{R} \int \rmdx{x_\text{t}} P_\text{t}(x_\text{t}) E(x_\text{d} | x_\text{t}) \,.
\end{equation}
\fred{In this section, all integration symbols are at the end, in ch. 2 they were right after the $\int$ sign.}
Here $P_\text{t}$ is the probability distribution of the true events,
the normalization constant $R$ is given by
\begin{equation}
  \label{eq:def-R}
    R \equiv \iint \rmdx{x_\text{t}} \rmdx{x_\text{d}} P_\text{t}(x_\text{t}) E(x_\text{d} | x_\text{t})\,.
\end{equation}
The kernel $E(x_\text{d}| x_\text{t})$ is usually decomposed as
\begin{equation}
  \Ekernel = \eps(\xt) P(\xd|\xt),
\end{equation}
where the PDF $P(\xd|\xt)$ models the resolution effects and the
unnormalized density $\eps(\xt)$ is the \emph{detector acceptance}
function. Perfect resolution corresponds to
\begin{equation}
  P(\xd|\xt) = \delta(\xd - \xt) \,.
\end{equation}

>>>>

Our initial motivation for studying what we wish to call the
\emph{method of moments} is the determination of angular observables
in the rare FCNC-mediated decay $\bar{B}\to \bar{K}^*(\to
\bar{K}\pi)\ell^+\ell^-$. However, the method we describe in the
following is general and applies to all decay or scattering
processes that can be formulated in terms of an orthogonal basis of angular functions.
We find a previous work \cite{Dighe:1998vk} that
advocates this method chiefly for the determination of angular
observables in non-leptonic $B$ decays but also mentions the
applicability to semileptonic decays. Our aim is to improve upon this
previous work by studying the uncertainties that are introduced by
mismodeling of the angular distribution, and by working out a recipe
to determine and unfold detector effects. The latter is crucial for
the application of the method to real data. We show that
the method of moments has several major advantages over the usual approach
based on likelihood fits:
\begin{enumerate}
    \item Likelihood fits have convergence problems for a small number of
        events, and can require reparametrizations and/or approximations for a
        successful fit to the signal PDF. As an example, see the LHCb
        analysis of the angular distribution in $\bar{B}\to
        \bar{K}^*\ell^+\ell^-$ decays \cite{Aaij:2013qta}.

        The method of
        moments does not require any such reparametrizations or approximations.

    \item Likelihood fits  can be unstable in case the underlying
        physical model is only partially known. This can lead to
        overestimating the number of physical parameters, and consequently
        inhibits the convergence of the fits. As an example, this type
        of problem occured in toy studies of the decay $B\to K^*(\to K\pi)\ell^+\ell^-$
        as reported in \cite{Egede:1142152}. It was subsequently solved
        when a missing symmetry relation between the angular observables
        was found and applied, thereby reducing the number of fit parameters.

        In contrast, we will show that the method of moments does not require
        information on the correlations between model parameters as an input.
        Instead, it yields the correlations as an output, at the expense of
        somewhat larger uncertainties.

    \item Mismodeling the underlying physics model can result in systematic bias
        in likelihood fits.

        We will
        show that the method of moments is insensitive to a certain type of mismodeling;
        i.e. introducing a cutoff in a partial-wave expansion of the signal PDF.

@@

We continue with basic definitions that pertain to angular observables, and our
results in the subsequent sections. Let $\thvec$ denote the set of all
angles, and let $\nuvec$ denote the set of all
other non-angular kinematic variables needed to fully specify the
final state of the process under study. For example, $\nuvec$ may
include invariant masses or center-of-mass energies. We define an
angular observable $S_i$ as a coefficient in the probability density
function (PDF), $P(\nuvec, \thvec)$, of the process by
\begin{align}
    \label{eq:def-P}
    P(\nuvec, \thvec) \equiv \sum_i S_i(\nuvec) \times f_i(\thvec)\,.
\end{align}
Here, the dependence on the decay angles $\thvec$ has been
explicitly factored out in terms of the angular functions
$\{f_i(\thvec)\}$. We assume there exists a dual basis of functions
$\{\dual{f}_i(\thvec)\}$ such that the orthonormality relations
\begin{equation}
    \label{eq:def-ortho-rel}
    \int_\Omega \rmdx{\vec{\theta}} \dual{f}_i(\thvec) f_j(\thvec)  = \delta_{ij}
\end{equation}
hold with $\Omega$ representing the full angular phase space relevant
to the process.  For particle decays, $P$ is generally expressed in
terms of the fully differential decay width,
\begin{align}
    \label{eq:def-P-decay}
    P(\nuvec, \thvec) \equiv \frac{1}{\Gamma}\frac{\rmdx{^{2}\Gamma}}{\rmdx{\vec\nu} \rmdx{\vec\theta}}\,,
\end{align}
where $\Gamma$ is the total decay width. For a scattering process, one can similarly use
\begin{align}
    \label{eq:def-P-scattering}
    P(\nuvec, \thvec) \equiv \frac{1}{\sigma}\frac{\rmdx{^{2}\sigma}}{\rmdx{\vec\nu} \rmdx{\vec\theta}}\,,
\end{align}
where the total cross section $\sigma$ is used for the
normalization. Since the determination of the total decay width or
total cross section can be quite difficult, we emphasize that
different normalizations for $P$ can be used.  For instance, the total
decay width (or cross section) of the process of interest can be
replaced by the corresponding quantity of a control-channel
process. This change of normalization is equivalent to a linear rescaling of
the angular observables $\lbrace S_i\rbrace$; thus ratios or similar suitable combinations
of the angular observables are not affected by a change of normalization.\\

Our method is an extension of the classical method of moments with
orthogonal functions~\cite[sec. 8.2]{James:2006zz}. The only difference
is that conventionally the angular functions are assumed
\emph{self-dual}, $\dual{f}_i = f_i$. However, it suffices that the
system of angular functions $\{f_i(\thvec)\}$ \emph{can} be transformed into
an orthonormal basis. We find it convenient to work
in the basis of Legendre polynomials that are not self-dual.
Our approach covers the self-dual case, provided that one
replace $\dual{f}_i \to f_i$ appropriately.
Using the ansatz
\begin{equation}
  \label{eq:dual-ansatz}
  \dual{f}_i = \sum_{j} a_{ij} f_j \,,
\end{equation}
the dual basis needs to be worked out case by case through solving the
linear system of equations (\ref{eq:def-ortho-rel}). For a selection
of hadron decays with a $b$ quark in the initial state and two leptons
in the final state, we list the dual bases in a series of appendices
\ref{app:btokll} through \ref{app:btokstarll}. Note that a
similar analysis was done in~\cite{Dighe:1998vk} for the decays $B \to
J/\psi \phi$ and $B \to J/\psi K^{*}$.

@@

Integration over the non-angular variables yields
\begin{equation}
    \langle S_i\rangle
    \equiv \int \rmdx{\vec\nu} S_i(\nuvec)
    = \int \rmdx{\vec \nu} \left[\int_{\Omega} \rmdx{\vec \theta} P(\nuvec,\thvec) \dual{f}_i(\thvec) \right].
\end{equation}
The remainder of this section describes the method of moments, in
which we replace the analytical integration by Monte Carlo (MC)
estimates.  The central tenet of MC integration is the fact that the
expectation value $E_P[g]$ of some function $g(x)$ under the
probability density $P(x)$,
\begin{equation}
    E_P[g] \equiv \int \rmdx{x} P(x) g(x),
\end{equation}
can be approximated by the consistent and unbiased
estimator $\est{E_P[g]}$~\cite[sec. 8.2]{James:2006zz}
\begin{equation}
    \label{eq:mc-id}
    E_P[g] \to \widehat{E_P[g]} \equiv \frac{1}{N} \sum_{n=1}^N g(x^{(n)}) \,,\,    x^{(n)} \sim P
\end{equation}
due to the strong law of large numbers for $N \to \infty$, assuming
that the variates $x^{(n)}$, $n = 1, \dots, N$, are distributed as
$P$.
Throughout this letter we denote all MC estimators with a wide hat.\\

@@

\item The estimate is unbiased for any $N$. In contrast, the
  maximum-likelihood estimate has a bias of order
  $1/N$~\cite{Cox:1968}. In practice, one should keep in mind the
  bias-variance trade-off: it is a well known phenomenon that removing
  the bias usually leads to an increase in variance of the sampling
  distribution of the estimator and vice versa~\cite[sec. 7.3]{James:2006zz}. From a
  Bayesian decision-theory point of view, both contribute similarly to
  the expected loss associated with deciding on just one value of the
  unknown parameter. One should therefore not prefer the method of moments
  over likelihood fits just because the former reduces
  the bias~\cite[sections 13.8,17.2]{jaynes:2003}. In fact, for
  the results discussed below in \refsec{numerics}, the likelihood
  fits --- if they converge --- exhibit a negligible bias and
  produce uncertainties $10\%$--$30\%$ smaller than those from the method of moments.

\item The approximate multivariate Gaussian distribution of $\vecest{S}$
    allows easier and more \add{precise} transfer of the information in the
  data to interested theorists for more accurate fits of
  standard-model and new-physics parameters \cite{Altmannshofer:2013foa,
Descotes-Genon:2013wba,Beaujean:2013soa},
  or for more precise predictions of optimized observables; see e.g.,
  \cite{Egede:2008uy,Egede:2010zc,Bobeth:2010wg,Becirevic:2011bp,
    Bobeth:2012vn,Matias:2012xw,DescotesGenon:2012zf} for definitions
  of such optimized observables in $B\to K^*\ell^+\ell^-$ decays,
  \cite{Faller:2013dwa} for application to the decay $B\to
  \pi\pi\ell^-\bar\nu_\ell$, and \cite{Boeer:2014xx} for observables
  in $\Lambda_b\to\Lambda(\to N\pi)\ell^+\ell^-$). While the
  likelihood also approaches a multivariate Gaussian as $N \to
  \infty$, the two methods differ in their utility as input for
  theorists if $\widehat{\vec{S}}$ is not well
  inside the physical region. For example, suppose there are two
  angular observables that are constrained to a triangular region by
  phase-space or unitarity arguments as
  \begin{equation}
    \label{eq:constr-ex}
    |S_1| + S_2 \le 1, \, S_1 \in [-1,1],\, S_2 \in [0,1] \,.
  \end{equation}
  It may (and often does) happen in practice that $\widehat{\vec{S}}$
  is close or even outside the allowed region such that a significant
  part of the probability mass covers unphysical values. In a Bayesian
  fit, one would take $\mathcal{N}(\vecest{S} | \vec{S},
  \widehat{\Sigma})$ as the sampling distribution of the ``data'', and
  simply set a uniform prior on the triangle in the $\vec{S}$ plane
  defined by \refeq{constr-ex} to have a well defined problem. This
  could be trivially combined with other independent information in a
  global fit. Someone with a different physics model might have to
  consider a
  different physical region, and could incorporate it just as easily.\\

  For a likelihood fit, the constraint needs to be part of the analysis
  performed by the experimental collaboration, and the resulting likelihood as
  a function of $\vec{S}$ may be distinctly not Gaussian.  Communicating such a
  result has proved to be challenging due to technical reasons (such as data
  formats, size etc.) This leads to the undesirable situation that only the
  mode and standard errors are reported, and theorists often include the
  results as independent measurements with a Gaussian distribution and
  disregard the boundary problem as well as correlations altogether.
\end{enumerate}

@@

For the selected class of decays, we can describe that problem as follows: The
PDF $P$ has a fixed dependence on the dilepton helicity angle $\theta_1$ and
the azimuthal angle $\theta_3$. (See also \refapp{btokstarll} for details on
the angular distribution.) However, at the level of decay amplitudes the
dimeson system can have an arbitrarily large total angular momentum $j$; only
its third component is restricted to $j_z = -1,0,+1$.  It is then convenient to
compute the angular distribution explicitly in terms of $\theta_1$ and
$\theta_3$, but leave the angular observables dependent on the remaining
helicity angle $\theta_2$ of the dimeson:
\begin{multline}
    P(\cos\theta_1,\cos\theta_2,\theta_3)\\
        \equiv \sum_k S_k(\nuvec,\cos\theta_2) f_k(\cos\theta_1,\theta_3)\,.
\end{multline}
We advocate here that is a sensible procedure to
perform an expansion in terms of Legendre polynomials $p_{l}^{(|m|)}$ with
respect to the remaining angle $\theta_2$. Here, the angular-momentum indices
$l$ and $m$ follow from the usual rules for addition of the angular momenta $j$
and $\tilde{j}$ of the partial-wave expansion of the underlying amplitude and
its complex conjugate: $j - \tilde{j} \leq l \leq j + \tilde{j}$, and $m = j_z
+ \tilde{j}_z$.

For the decay at hand, we consider the partial-wave expansion for the angular observables (see \cite{Lee:1992ih} and \refapp{partial-waves})
\begin{equation}
    S_k(\vec{\nu}) \equiv \frac{I_k(\nuvec)}{4\pi \Gamma}\,,\quad\text{with}\quad \Gamma = I_1(\nuvec) - \frac{I_2(\nuvec)}{3}\,,
\end{equation}
which reads
\begin{multline}
    \label{eq:expansion}
    S_{k}(\vec{\nu},\cos\theta_2) \\
        \equiv \sum_{l=0}^\infty \frac{1}{n_{l,|m|}} S_{k,l}(\vec{\nu}) p_{l}^{(|m|)}(\cos\theta_2)\,.
\end{multline}
The normalization factor $n_{l,|m|}$ is defined in \refeq{legendre-scalar-product}. Within our
example we have (cf. also \refapp{btokstarll} and \cite{Lee:1992ih})
\begin{equation}
    |m| = \begin{cases}
        0\,, & k = 1,2,6\\
        1\,, & k = 4,5,7,8\\
        2\,, & k = 3,9
    \end{cases}\,.
\end{equation}

The angular observables $S_{k,l}$ -- as defined in \refeq{expansion} -- have
the merit of a well defined total angular momentum, and thus are physically
distinguishable.  As a consequence of the orthogonality of the Legendre
polynomials, any mismodeling (or rather, lack of modelling) of higher
partial-wave observables \emph{does not affect} the method of moments as
discussed in the previous section. That is to say, adding further (orthogonal) terms
to the PDF only appends observables to $\vec{S}$, but does not change the leading
elements. The same applies to the covariance.\\

@@

Ascertaining a detector's performance to detect signal events
with accurate determination of the event's angles is generally a
difficult task. In the ideal case, one would have an explicit
probabilistic model of the detector acceptance and could thus
write down the full forward model from which the measured events
arise. In practice, that is not feasible, and one is
forced to simplify the model.
The standard approach is to generate the \emph{true} particle events
$\lbrace x_\text{t}^{(n)}\rbrace =
\lbrace(\vec{\nu}^{(n)}_\text{t},\vec\theta^{(n)}_\text{t})\rbrace$,
$n=1,\dots,N_\text{t}$, from a PDF assumed to describe the bare
physical process, and to propagate those particles through a detailed
simulation of the detector.  The observable traces that the particles
leave in the detector are fed into reconstruction algorithms resulting
in the \emph{detector} events $\lbrace x^{(n)}_\text{d}\rbrace =
\lbrace(\vec{\nu}^{(n)}_\text{d}, \vec\theta^{(n)}_\text{d})\rbrace$,
$n=1,\dots,N_\text{d}$ with $N_\text{d} \le N_\text{t}$.  In general,
the distribution of the detected events is
\begin{equation}
    P_\text{d}(x_\text{d}) = \frac{1}{R} \int \rmdx{x_\text{t}} P_\text{t}(x_\text{t}) E(x_\text{d} | x_\text{t}) \,.
\end{equation}
Here $P_\text{t}$ is the probability distribution of the true events,
the normalization constant $R$ is given by
\begin{equation}
  \label{eq:def-R}
    R \equiv \iint \rmdx{x_\text{t}} \rmdx{x_\text{d}} P_\text{t}(x_\text{t}) E(x_\text{d} | x_\text{t})\,.
\end{equation}
The kernel $E(x_\text{d}| x_\text{t})$ is usually decomposed as
\begin{equation}
  \Ekernel = \eps(\xt) P(\xd|\xt),
\end{equation}
where the PDF $P(\xd|\xt)$ models the resolution effects and the
unnormalized density $\eps(\xt)$ is the \emph{detector acceptance}
function. Perfect resolution corresponds to
\begin{equation}
  P(\xd|\xt) = \delta(\xd - \xt) \,.
\end{equation}

In what follows, we propose a systematic method to unfold all effects
of $\Ekernel$ through MC simulations and the method of moments,
using that $\Ekernel$ can be expanded --- at
least formally --- in Legendre polymials.  For illustration, we
proceed with the explicit example of a uniangular\footnote{%
  The generalization of this section to multiangular PDFs is
  straighforward. It can be achieved by promoting $x$ to a vector,
  promoting the Legendre polynomials to products of independent
  polynomials or spherical harmonics, and promoting the indices
  $i,j,k,m$ to multi-indices.  } PDF with $x = \cos\theta$. Let us
define the PDF in terms of the Legendre polymials (i.e., $f_k(x)
\equiv p_k(x)$) and angular observables $\vec{S}$ as
\begin{equation}
    P\subt(\xt) \equiv P\subt(\xt | \vec{S}) = \sum_k S_k f_k(\xt),
  \end{equation}
where $k = 0, 1, \dots$ denotes an angular-momentum-like index associated with the observables.
Normalization of $P\subt$ is equivalent to choosing $S_0 = 1/2$. Requiring $P\subt(\xt) \ge 0 \,\forall \xt$
implies $|S_k| \leq 1/2$ for $k > 0$. More stringent relations between the
$S_k$ might hold, but are of no concern here. For later use, we define
\begin{equation}
  \label{eq:def-Sk}
  S^{(m)}_k \equiv
  \begin{cases}
    1/2 \delta_{k,0}, & m=0\\
    1/2 (\delta_{k,0} + \delta_{k,m}), & m>0 \,,
  \end{cases}
\end{equation}
and note that
\begin{equation}
    P^{(m)}(\xt) \equiv P(\xt | \vec{S}^{(m)})
\end{equation}
is a valid PDF. The dual basis of angular functions follows then from the normalization
and orthogonality of the Legendre polynomials, and one therefore has
\begin{gather}
    \tilde{f}_k(x) = \frac{2 k + 1}{2} f_k(x)\,,\\
    \int_{-1}^{+1} \rmdx{x} \tilde{f}_k(x) f_l(x) = \delta_{k,l}\,.
\end{gather}
We now define the \emph{simulated raw} moments $\vec{Q}^{(m)}$ as
\begin{equation}
    \label{eq:analytic-raw-moments}
    Q_i^{(m)} \equiv \iint \rmdx{\xt} \rmdx{\xd} \tilde{f}_i(\xd) P^{(m)}(\xt) \Ekernel,
\end{equation}
which are instrumental to our recipe. Monte Carlo estimators of these moments can be constructed from specifically crafted detector events $x_\text{d}^{(n,m)}$, $n = 1, \dots, N_\text{d}^{(m)}$, where
\begin{equation}
    x_\text{d}^{(n,m)} \sim P_E^{(m)}(\xd) \equiv \frac{1}{R^{(m)}} \int \rmdx{\xt} P^{(m)}(\xt) \Ekernel\,.
\end{equation}
In words, for each $m$ it is required to generate events from
a toy physical distribution $P^{(m)}$, for which $S_0 = S_m=1/2$ and all other
observables are set to zero. Next, propagate these events through
a detector simulation. The normalization $R^{(m)}$ is chosen such
that $\int P_E^{(m)}(\xd) \rmdx{\xd} = 1$.  We emphasize that
$R^{(m)}$ can be estimated as $\est{R}^{(m)} = N_\text{d}^{(m)} /
N_\text{t}$, where $N_\text{t}$ corresponds to the number of
simulated true events.  The estimators then read
\begin{equation}
    \est{Q}_i^{(m)} \equiv \est{R}^{(m)} \frac{1}{N_\text{d}} \sum_n^{N_\text{d}} \tilde{f}_i(x_\text{d}^{(n,m)})
\end{equation}

Linearity of the integral over $x$ and convergence of the expansion of
$P_E^{(m)}$ in terms of Legendre polynomials ensures that
\begin{equation}
    \vec{Q}^{(m)} = M \vec{S}^{(m)}\,.
\end{equation}
We call the matrix $M^{-1}$ the \emph{unfolding matrix},
which is specific to the decay at hand.
Given our definition of $S_k^{(m)}$ in
\refeq{def-Sk} it is easy to see that
\begin{equation}
    M_{ij} = \begin{cases}
        2 Q_i^{(0)}                          & j = 0\,,\\
        2\left(Q_i^{(j)} - Q_i^{(0)}\right)  & j \neq 0\,,\\
    \end{cases}
\end{equation}
and its MC estimator $\est{M}$  can be obtained through the replacements $Q_i^{(m)} \to \est{Q}_i^{(m)}$.\\
